\vspace{-\baselineskip} Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-2\baselineskip}

### 2017-06-02 paper/ms.tex
Minor edits of simulation section.

Simulation of a Gaussian process for large $N$ requires multiplying an
$N$-dimensional vector of Gaussian normal deviates, $\bf{q}$, by the square root
of the covariance matrix:
\begin{equation}
{\bf y} = K_\bvec{\alpha}^{1/2} {\bf q},
\end{equation}
where $q_i \sim N(0,1)$.  Note that $K_\bvec{\alpha} = K_\bvec{\alpha}^{1/2} \left(K_\bvec{\alpha}^{1/2}\right)^T$, which
comparing with the Cholesky factorization gives $K_\bvec{\alpha}^{1/2} = L D^{1/2}$,
where $(D^{1/2})_{ii} = \sqrt{D_{ii}}$.  Also note that the exponent of
the likelihood with $\mu_\bvec{\theta} =0$ is given by
\begin{eqnarray}
Q &=&-\frac{1}{2}{\bf y}^T K_\bvec{\alpha}^{-1} {\bf y}\\
&=& -\frac{1}{2}{\bf y}^T \left(K_\bvec{\alpha}^{-1/2}\right)^T K_\bvec{\alpha}^{-1/2} {\bf y}\\
&=& -\frac{1}{2}{\bf q}^T K_\bvec{\alpha}^{1/2} \left(K_\bvec{\alpha}^{-1/2}\right)^T K_\bvec{\alpha}^{-1/2} K_\bvec{\alpha}^{1/2} {\bf q}\\
&=& -\frac{1}{2}{\bf q}^T {\bf q}.
\end{eqnarray}
The last quantity is simply the dot product of an $N$-dimensional vector of independent standard normal random variables with itself,
which is distributed according to the chi-squared distribution with $N$ degrees of freedom, i.e.\ $Q \sim \chi_N^2$.

>>>>

Simulation of a Gaussian process for large $N$ requires multiplying an
$N$-dimensional vector of Gaussian normal deviates, $\bf{q}$, by the square root
of the covariance matrix:
\begin{equation}
{\bf y} = K_\bvec{\alpha}^{1/2} {\bf q},
\end{equation}
where $q_i \sim N(0,1)$.  Note that $K_\bvec{\alpha} = K_\bvec{\alpha}^{1/2} \left(K_\bvec{\alpha}^{1/2}\right)^T$, which
comparing with the Cholesky factorization gives $K_\bvec{\alpha}^{1/2} = L D^{1/2}$,
where $(D^{1/2})_{ii} = \sqrt{D_{ii}}$.  Also note that the first term in $-2\ln{\mathcal{L}}$
with $\mu_\bvec{\theta} =0$ is given by
\begin{eqnarray}
\mathcal{Q} &=&{\bf y}^T K_\bvec{\alpha}^{-1} {\bf y}\\
&=& {\bf y}^T \left(K_\bvec{\alpha}^{-1/2}\right)^T K_\bvec{\alpha}^{-1/2} {\bf y}\\
&=& {\bf q}^T K_\bvec{\alpha}^{1/2} \left(K_\bvec{\alpha}^{-1/2}\right)^T K_\bvec{\alpha}^{-1/2} K_\bvec{\alpha}^{1/2} {\bf q}\\
&=& {\bf q}^T {\bf q}.
\end{eqnarray}
This last quantity is simply the dot product of an $N$-dimensional vector of independent standard normal random variables with itself,
which is distributed according to the chi-squared distribution with $N$ degrees of freedom, i.e.\ $\mathcal{Q} \sim \chi_N^2$.
